Self Referencing Sequences

### 2014-10-27 draft.tex
edited up to theorem 1, which I changed to lemma 1.

\subsection{Problem Setup}
Consider the sequence
\begin{align*}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{align*}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{align*}
[1],[2,2],[1,1],[2],[1],[2,2],[1],[2,2],[1,1],[2],[1,1],[2,2],\dots
\end{align*}
The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}, and for the remainder of this paper we study properties of such sequences.

We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so the above sequence is generated by $\{1,2\}$. We will refer to this set as the \emph{generating set}. Observe that it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. However, in a sequence with generating set $\{1,2,3\}$, for example, a block of 1's could be followed by either a block of 2's or a block of 3's. Thus, we will assume that the next number to use is the following element (in cyclic order) in the generating set. (Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.)

@@

\section{Prefixes}
\label{prefixes}
\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
\begin{equation*}
a_{n+i} = b_{i},
\end{equation*}
for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
\end{definition}
Note that a pair of sequences can have prefixes of different length. For example if we have 
\begin{align*}
&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
\end{align*}
 then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
 \begin{definition} 
 We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
 \end{definition}
We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$.

\begin{theorem} The self-referencing sequence beginning with a $1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ differs from the self-referencing sequence beginning with $c_1$ generated over $\{1, c_1, c_2, \ldots, c_n\}$ by a minimal prefix of length 1.
\end{theorem}
\begin{proof}
Let $a_i$ refer to the $i$th term of the sequence beginning with a $1$ and $b_i$ refer to the $i$th term of the sequence beginning with $c_1$. We know that $a_1=1$, this means that the first block in $a_i$ must be of length 1. In other words we must have $a_2=c_1$, but since $a_2$ is now part of the second block of $a_i$ the value of $a_3$ is not specified by $a_1$. Thus we could specify the rest of $a_i$ only knowing the value of $a_2$, but $b_1=c_1$, so this is exactly how the values of $b_i$ is specified. Therefore
\begin{equation*}
a_{1+i} = b_{i},
\end{equation*}
so $a_i$ and $b_i$ differ by a prefix of length 1, but $a_{1} \neq b_{1}$ so it is a minimal prefix. 
\end{proof}

>>>>

\subsection{Problem Setup}
Consider the sequence
\begin{align*}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{align*}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{align*}
[1],[2,2],[1,1],[2],[1],[2,2],[1],[2,2],[1,1],[2],[1,1],[2,2],\dots
\end{align*}
The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}. For the remainder of this paper we study properties of such sequences.

We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so the above sequence is generated by $\{1,2\}$. We will refer to this set as the \emph{generating set}. Observe that it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. There does exist ambiguity in a sequence with generating set $\{1,2,3\}$, since after a block of 1's, ther can be a block of 2's or a block of 3's. We assign an order to the generating set and resolve the ambiguity. The order we will use will be the order of the elements presented, for example, $\{1, 2, 3 \}$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$. Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.

@@

\section{Prefixes}
\label{prefixes}
\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
\begin{equation*}
a_{n+i} = b_{i},
\end{equation*}
for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
\end{definition}
Note that a pair of sequences can have prefixes of different length. For example if we have 
\begin{align*}
&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
\end{align*}
 then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
 \begin{definition} 
 We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
 \end{definition}
We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$. The following lemma is a trivial consequence of our algorithm for generating self-referential sequences:

\begin{lemma} Let $C = \{1, c_1, c_2, \ldots, c_n\}$, $s(C, 1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
\end{lemma}
\begin{proof}
Let $a_i$ refer to the $i$th term of $s(C, 1)$ and $b_i$ refer to the $i$th term of $s(C, c_1)$. If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = c_1$, and that this is the beginning of a new block of length $c_1$. Furthermore, $b_1 = c_1$ and it indicated the beginning of a block of length $c_1$. Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements. Because the number and the begining of the block is the same. The algorithm will generate the same sequence, shifted by 1. Therefore, $s(C,1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
\end{proof}
